---
title: "The Impact of Russia Ukraine conflict on the oil Prices in Europe."
author: "Farman Karim"
date: "2024-01-19"
output: html_document
---

```{r include=FALSE}
library(tidyr)
library(dplyr)
library(tidyselect)
library(readxl)
library(plm)
library(ggplot2)
library(tidyverse)
install.packages('readxl')
```

```{r setup, include=FALSE}
Diesel_data <- read_excel("/home/faru/R/project/Diesel_data.xlsx")
View(Diesel_data)
names(Diesel_data)
```

```{r}
names(Diesel_data)[names(Diesel_data) == "VALUE"] <- "PRICE"
names(Diesel_data)[names(Diesel_data) == "TIME"] <- "DATE"
```

```{r}
Diesel_data <- subset(Diesel_data, select = -c(PRODUCT, FLOW, UNIT))
colnames(Diesel_data)
```

```{r}
Diesel_data$DATE <- as.Date(Diesel_data$DATE)
Diesel_data$PRICE <- as.numeric(gsub("[^0-9.]", "", Diesel_data$PRICE))
war_start_date <- as.Date("2022-02-01")


```

Diesel_data\$DATE \<- as.Date(Diesel_data\$DATE)

```{r}
Diesel_data$Log_price = log(Diesel_data$PRICE)
Diesel_data$Log_price <- as.numeric(gsub("[^0-9.]", "", Diesel_data$Log_price))
colnames(Diesel_data)
```

```{r}
russia_neighbors <- c("Estonia", "Finland", "Hungary", "Lithuania", "Poland", "Slovak Republic")
Diesel_data$NEIGHBORS = ifelse (Diesel_data$COUNTRY %in% russia_neighbors, 1, 0)

```

```{r}
Diesel_data$INVASION = ifelse (war_start_date > Diesel_data$DATE, 0, 1)
```

```{r}

# interaction dummy variable 
Diesel_data$INTERACTION <- Diesel_data$NEIGHBORS * Diesel_data$INVASION
names(Diesel_data)
```

```{r}
before_invasion_data <- Diesel_data[Diesel_data$INVASION == 0, ]
summary(before_invasion_data$Log_price)
```

```{r}
after_invasion_data <- Diesel_data[Diesel_data$INVASION == 1, ]
summary(after_invasion_data$Log_price)

```

```{r}
by(Diesel_data$Log_price, list(Diesel_data$INVASION, Diesel_data$NEIGHBORS), summary)

```

```{r}
ggplot(Diesel_data, aes(x = DATE, y = Log_price, color = COUNTRY)) +
  geom_line() +
  scale_x_date(date_breaks = "12 months", date_labels = "%b, %Y") +
  facet_wrap(~COUNTRY, scales = "free_y", ncol = 4) +
  labs(title = "Oil Prices Before and After Russia-Ukraine War",
       x = "Date", y = "Oil Price") +
  theme_minimal()
```

```{r}
Diesel_data %>%
  ggplot(aes(x = DATE, y = Log_price, color = COUNTRY)) +
    labs(x = "Date", y = "Oil Prices")+
  geom_line() +
  theme_minimal()

```

```{r}
selected_countries <- c("Netherlands ", "Belgium", "Portugal", "Luxembourg ", "France", 
                        "Germany","Estonia", "Finland", "Lithuania","Hungary","Poland", 
                        "Slovak Republic")

filtered_data <- Diesel_data[Diesel_data$COUNTRY %in% selected_countries, ]

```

```{r}
hist(filtered_data$Log_price[filtered_data$INVASION == 0], main = "Distribution of 'VALUE' Before Invasion", xlab = "VALUE", col = "white", breaks = 30)
abline(v = mean(filtered_data$Log_price[filtered_data$INVASION == 0]), col = "red", lwd = 2)

```

```{r}
hist(filtered_data$Log_price[filtered_data$INVASION == 1], main = "Distribution of 'VALUE' After Invasion", xlab = "VALUE", col = "grey", breaks = 30)
abline(v = mean(filtered_data$Log_price[filtered_data$INVASION == 1]), col = "red", lwd = 2)

```

```{r}
t.test(before_invasion_data$Log_price, after_invasion_data$Log_price)
```

```{r}
t.test(filtered_data$Log_price[filtered_data$NEIGHBORS == 1], filtered_data$Log_price[filtered_data$NEIGHBORS == 0], var.equal = TRUE)

```

```{r}
log_value_of_price = filtered_data$Log_price
ggplot(filtered_data, aes(x = factor(INVASION), y = log_value_of_price, fill = factor(NEIGHBORS))) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(x = "Invasion Status", y = "Mean Oil Prices", fill = "Neighbors") +
  ggtitle("Mean Oil Prices Before and After Invasion for Neighboring and Non-Neighboring Countries")

```

```{r}
ggplot(filtered_data, aes(x = factor(INVASION), y = log_value_of_price, fill = factor(NEIGHBORS))) +
  geom_boxplot() +
  labs(x = "Invasion Status", y = "Oil Prices", fill = "Neighbors") +
  ggtitle("Distribution of Oil Prices Before and After Invasion for Neighboring and Non-Neighboring Countries")
```

```{r}
ggplot(filtered_data, aes(x = DATE, y = log_value_of_price, color = factor(NEIGHBORS))) +
  geom_point() +
  labs(x = "Date", y = "Oil Prices", color = "Neighbors") +
  ggtitle("Scatter Plot of Oil Prices Over DATE for Neighboring and Non-Neighboring Countries")

```

```{r}
plot_data <- filtered_data %>% 
  # Make these categories instead of 0/1 numbers so they look nicer in the plot
  mutate(INVASION = factor(INVASION, labels = c("Before WAR", "After WAR")),
         NEIGHBORS = factor(NEIGHBORS, labels = c("NON-NEIGHBORS", "NEIGHBORS"))) %>% 
  group_by(NEIGHBORS, INVASION) %>% 
  summarize(mean_duration = mean(PRICE),
            se_duration = sd(PRICE) / sqrt(n()),
            upper = mean_duration + (1.96 * se_duration),
            lower = mean_duration + (-1.96 * se_duration)) 

ggplot(plot_data, aes(x = INVASION, y = mean_duration)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  color = "darkgreen", size = 1) +
  facet_wrap(vars(NEIGHBORS))
```

```{r}
ggplot(plot_data, aes(x = INVASION, y = mean_duration, color = NEIGHBORS)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), size = 1) + 
  geom_line(aes(group = NEIGHBORS))
```
